Generalized Plackett-Luce Likelihoods

### 2023-05-16 inst/jss_hyper3.Rnw
fixes #205

\subsubsection{Fisher information}
If we have two clones of $a$ and a singleton $b$, then there are three
possible rank statistics: (i), $a\succ a\succ b$ with probability
$\frac{a^2}{1+a}$; (ii), $a\succ b\succ a$, with
$\frac{1-2a-a^2}{(1+a)^2}$, (iii), $b\succ a\succ a$ at
$\frac{1-a}{1+a}$.  Likelihood functions for these order statistics
are given in Figure~\ref{ABAetc_likelihood}.  It can be shown that the
Fisher information for such observations is

\begin{equation}
\mathcal{I}(a)=\frac{1+16a+18a^2-4a^3-15a^4-7a^5-a^6}{a(1-a)(1+a)^3}
\end{equation}

which has a minimum of about $13.9$ at at about $a=0.575$.  We can
compare this with the Fisher information matrix ${\mathcal I}$, for
the case of three distinct runners $a,b,c$, evaluated at its minimum
of $p_a=p_b=p_c=\frac{1}{3}$.  If we observe the complete order
statistic, $\left|{\mathcal I}\right| =\frac{1323}{16}\simeq 82.7$; if
we observe just the winner, $\left|{\mathcal I}\right|=27$, and if we
observe just the loser we have $\left|{\mathcal
I}\right|=\frac{16875}{256}\simeq 65.9$.  A brief discussion is given
at \code{inst/fisher\_inf\_PL3.Rmd}.

\begin{figure}[htbp]
  \begin{center}
<<figabbabb,echo=FALSE,fig=TRUE>>=
f_aab <- function(a){a^2 / (1 + a)}
f_aba <- function(a){a * (1 - a) / (1 + a)^2}
f_baa <- function(a){(1 - a) / (1 + a)}
p <- function(f, ...){
  a <- seq(from = 0, by = 0.005, to = 1)
  points(a, f(a) / max(f(a)), ...)
  }
plot(0:1, 0:1, xlab = expression(p[a]), ylab = "Likelihood", type = "n")
p(f_aab, type = "l", col = "black")
p(f_aba, type = "l", col = "red")
p(f_baa, type = "l", col = "blue")
text(0.8,0.8,"AAB")
text(0.8,0.5,"ABA",col="red")
text(0.8,0.15,"BAA",col="blue")
abline(h = exp(-2), lty = 2)
@
\caption{Likelihood \label{ABAetc_likelihood} functions for
observations $a\succ a\succ b$, $a\succ b\succ a$, $b\succ a\succ a$.  Horizontal dotted
line represents two units of support}
\end{center}
\end{figure}

>>>>

\subsubsection{Fisher information}
If we have two clones of $a$ and a singleton $b$, then there are three
possible rank statistics: (i), $a\succ a\succ b$ with probability
$\frac{2a^2}{1+a}$; (ii), $a\succ b\succ a$, with
$\frac{2a(1-a)}{(1+a)}$, (iii), $b\succ a\succ a$ at
$\frac{1-a}{1+a}$.  Likelihood functions for these order statistics
are given in Figure~\ref{ABAetc_likelihood}.  It can be shown that the
Fisher information for such observations is

\begin{equation}
\mathcal{I}(a)=2\frac{1+a+a^2}{a(1-a)(1+a)^2}
\end{equation}

which has a minimum of about $6.21$ at at about $a=0.522$.  We can
compare this with the Fisher information matrix ${\mathcal I}$, for
the case of three distinct runners $a,b,c$, evaluated at its minimum
of $p_a=p_b=p_c=\frac{1}{3}$.  If we observe the complete order
statistic, $\left|{\mathcal I}\right| =\frac{1323}{16}\simeq 82.7$; if
we observe just the winner, $\left|{\mathcal I}\right|=27$, and if we
observe just the loser we have $\left|{\mathcal
I}\right|=\frac{16875}{256}\simeq 65.9$.  A brief discussion is given
at \code{inst/fisher\_inf\_PL3.Rmd}.

\begin{figure}[htbp]
  \begin{center}
<<figabbabb,echo=FALSE,fig=TRUE>>=
f_aab <- function(a){a^2 / (1 + a)}
f_aba <- function(a){a * (1 - a) / (1 + a)}
f_baa <- function(a){(1 - a) / (1 + a)}
p <- function(f, ...){
  a <- seq(from = 0, by = 0.005, to = 1)
  points(a, f(a) / max(f(a)), ...)
  }
plot(0:1, 0:1, xlab = expression(p[a]), ylab = "Likelihood", type = "n")
p(f_aab, type = "l", col = "black")
p(f_aba, type = "l", col = "red")
p(f_baa, type = "l", col = "blue")
text(0.8,0.8,"AAB")
text(0.8,0.4,"ABA",col="red")
text(0.8,0.05,"BAA",col="blue")
abline(h = exp(-2), lty = 2)
@
\caption{Likelihood \label{ABAetc_likelihood} functions for
observations $a\succ a\succ b$, $a\succ b\succ a$, $b\succ a\succ a$.  Horizontal dotted
line represents two units of support}
\end{center}
\end{figure}
